### 2019-09-13 ic-journal/multivariate.tex
Updated multivariate substitution theorem list

\begin{lemma}[\texttt{CCS_SUBST_SING}]
  If there is only one single variable $X$ in the map (with the target
  expression $E'$), then \HOLinline{\HOLConst{CCS_SUBST}} behaves exactly the
  same as (the univariate) \HOLinline{\HOLConst{CCS_Subst}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{CCS_SUBST} \ensuremath{(}\HOLConst{FEMPTY} \HOLSymConst{|+} \ensuremath{(}\HOLFreeVar{X}\HOLSymConst{,}\HOLFreeVar{E\sp{\prime}}\ensuremath{)}\ensuremath{)} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{CCS_Subst} \HOLFreeVar{E} \HOLFreeVar{E\sp{\prime}} \HOLFreeVar{X}
\end{alltt}
\end{lemma}

@@

@@

For instance, to see the weak guardedness of $a.X + b.X + c.Y$, it is
natural to focus on each (free) variable: if we substitute each
occurrence of just one variable, say $X$, with a ``hole'' and see the
resulting term as a $\lambda$-function, then all $\lambda$-functions
derived from the original expression in this way must satisfy
\HOLinline{\HOLConst{WG}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{Y}\ensuremath{)}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}
\end{alltt}
Notice that, \HOLinline{\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\\\;\HOLSymConst{\ensuremath{=}}\\\;\ensuremath{[}\HOLFreeVar{t}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\ensuremath{(}\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\ensuremath{)}}. This idea leads to the following definitions:
\begin{alltt}
   \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{CONTEXT} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}

   \HOLConst{weakly_guarded} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLSymConst{\HOLTokenDefEquality{}}
     \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}
\end{alltt}

@@

\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{CCS_equation} \HOLFreeVar{Xs} \HOLFreeVar{Es} \HOLSymConst{\HOLTokenEquiv{}}
   \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{FV} \HOLBoundVar{e} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLFreeVar{Es}

@@

\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{weakly_guarded} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{FV} \HOLFreeVar{E} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{set} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps}.
       \HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{u} \HOLBoundVar{P\sp{\prime}}.
           \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sp{\prime}}.
               \HOLConst{context} \HOLFreeVar{Xs} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{FV} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{set} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenConj{}}
               \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}. \HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}} \ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLBoundVar{E\sp{\prime}}

>>>>

@@

The next three lemmas show the relative correctness of \texttt{CCS\_SUBST}
w.r.t. \texttt{CCS\_Subst}:
\begin{lemma}[\texttt{CCS_SUBST_SING}]
  If there is only one single variable $X$ in the map (with the target
  expression $E'$), then \HOLinline{\HOLConst{CCS_SUBST}} behaves exactly the
  same as (the univariate) \HOLinline{\HOLConst{CCS_Subst}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{CCS_SUBST} \ensuremath{(}\HOLConst{FEMPTY} \HOLSymConst{|+} \ensuremath{(}\HOLFreeVar{X}\HOLSymConst{,}\HOLFreeVar{E\sp{\prime}}\ensuremath{)}\ensuremath{)} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{CCS_Subst} \HOLFreeVar{E} \HOLFreeVar{E\sp{\prime}} \HOLFreeVar{X}
\end{alltt}
\end{lemma}

\begin{lemma}[\texttt{CCS_SUBST_reduce}]
Under certain conditions (to get rid of substitution orders), a \multivariate substitution of variables
\HOLinline{\HOLFreeVar{X}\HOLSymConst{::}\HOLFreeVar{Xs}} can be reduced to a \multivariate substitution of
variables \HOLinline{\HOLFreeVar{Xs}} and an \univariate substitution of $X$:
\begin{alltt}
\HOLTokenTurnstile{} \HOLSymConst{\HOLTokenNeg{}}\HOLConst{MEM} \HOLFreeVar{X} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}}
   \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLFreeVar{X} \HOLSymConst{\HOLTokenNotIn{}} \HOLConst{FV} \HOLBoundVar{e}\ensuremath{)} \HOLFreeVar{Ps} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{E}. \ensuremath{[}\HOLFreeVar{P}\HOLSymConst{::}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{X}\HOLSymConst{::}\HOLFreeVar{Xs}\ensuremath{]} \HOLBoundVar{E} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLFreeVar{P}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]} \ensuremath{(}\ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLBoundVar{E}\ensuremath{)}    
\end{alltt}
\end{lemma}

\begin{lemma}[\texttt{CCS_SUBST_FOLDR}]
Under certain conditions (to get rid of substitution orders), a
\multivariate substitution can be reduced to repeated applications
of \univariate substitutions of each variable:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}}
   \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\ensuremath{(}\HOLBoundVar{x}\HOLSymConst{,}\HOLBoundVar{p}\ensuremath{)}. \HOLConst{FV} \HOLBoundVar{p} \HOLSymConst{\HOLTokenSubset{}} \HOLTokenLeftbrace{}\HOLBoundVar{x}\HOLTokenRightbrace{}\ensuremath{)} \ensuremath{(}\HOLConst{ZIP} \ensuremath{(}\HOLFreeVar{Xs}\HOLSymConst{,}\HOLFreeVar{Ps}\ensuremath{)}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \ensuremath{[}\HOLFreeVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLSymConst{\ensuremath{=}} \HOLConst{FOLDR} \ensuremath{(}\HOLTokenLambda{}\ensuremath{(}\HOLBoundVar{x}\HOLSymConst{,}\HOLBoundVar{y}\ensuremath{)} \HOLBoundVar{e}. \ensuremath{[}\HOLBoundVar{y}\ensuremath{/}\HOLBoundVar{x}\ensuremath{]} \HOLBoundVar{e}\ensuremath{)} \HOLFreeVar{E} \ensuremath{(}\HOLConst{ZIP} \ensuremath{(}\HOLFreeVar{Xs}\HOLSymConst{,}\HOLFreeVar{Ps}\ensuremath{)}\ensuremath{)}    
\end{alltt}
\end{lemma}

@@

For instance, to see the weak guardedness of $a.X + b.X + c.Y$, it is
natural to focus on each (free) variable: if we substitute each
occurrence of just one variable, say $X$, with a ``hole'' and see the
resulting term as a $\lambda$-function, then all $\lambda$-functions
derived from the original expression in this way must satisfy
\HOLinline{\HOLConst{WG}}:
\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{Y}\ensuremath{)}
\HOLTokenTurnstile{} \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var} \HOLFreeVar{X} \HOLSymConst{\ensuremath{+}} \HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLBoundVar{t}\ensuremath{)}
\end{alltt}
Notice that, \HOLinline{\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLFreeVar{t}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\\\;\HOLSymConst{\ensuremath{=}}\\\;\ensuremath{[}\HOLFreeVar{t}\ensuremath{/}\HOLFreeVar{X}\ensuremath{]}\\\;\ensuremath{(}\HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{b}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{X}\\\;\HOLSymConst{\ensuremath{+}}\\\;\HOLFreeVar{c}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{var}\\\;\HOLFreeVar{Y}\ensuremath{)}}. This idea leads to the following definitions:
\begin{alltt}
   \HOLConst{context} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{CONTEXT} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}

   \HOLConst{weakly_guarded} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{X}. \HOLConst{WG} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{t}. \ensuremath{[}\HOLBoundVar{t}\ensuremath{/}\HOLBoundVar{X}\ensuremath{]} \HOLFreeVar{E}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Xs}
\end{alltt}

@@

\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{CCS_equation} \HOLFreeVar{Xs} \HOLFreeVar{Es} \HOLSymConst{\HOLTokenEquiv{}}
   \HOLConst{ALL_DISTINCT} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{LENGTH} \HOLFreeVar{Es} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{FV} \HOLBoundVar{e} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLFreeVar{Es} \HOLSymConst{\HOLTokenConj{}}
   \HOLConst{EVERY} \ensuremath{(}\HOLTokenLambda{}\HOLBoundVar{e}. \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{e}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)}\ensuremath{)} \HOLFreeVar{Es}

@@

\begin{alltt}
\HOLTokenTurnstile{} \HOLConst{weakly_guarded} \HOLFreeVar{Xs} \HOLFreeVar{E} \HOLSymConst{\HOLTokenConj{}} \HOLConst{FV} \HOLFreeVar{E} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{set} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLFreeVar{E}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenImp{}}
   \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Ps}.
       \HOLConst{LENGTH} \HOLBoundVar{Ps} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}}
       \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{u} \HOLBoundVar{P\sp{\prime}}.
           \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\HOLTokenImp{}}
           \HOLSymConst{\HOLTokenExists{}}\HOLBoundVar{E\sp{\prime}}.
               \HOLConst{context} \HOLFreeVar{Xs} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenConj{}} \HOLConst{FV} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenSubset{}} \HOLConst{set} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenConj{}} \HOLConst{DISJOINT} \ensuremath{(}\HOLConst{BV} \HOLBoundVar{E\sp{\prime}}\ensuremath{)} \ensuremath{(}\HOLConst{set} \HOLFreeVar{Xs}\ensuremath{)} \HOLSymConst{\HOLTokenConj{}}
               \HOLBoundVar{P\sp{\prime}} \HOLSymConst{\ensuremath{=}} \ensuremath{[}\HOLBoundVar{Ps}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLBoundVar{E\sp{\prime}} \HOLSymConst{\HOLTokenConj{}}
               \HOLSymConst{\HOLTokenForall{}}\HOLBoundVar{Qs}. \HOLConst{LENGTH} \HOLBoundVar{Qs} \HOLSymConst{\ensuremath{=}} \HOLConst{LENGTH} \HOLFreeVar{Xs} \HOLSymConst{\HOLTokenImp{}} \ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLFreeVar{E} \HOLTokenTransBegin\HOLBoundVar{u}\HOLTokenTransEnd \ensuremath{[}\HOLBoundVar{Qs}\ensuremath{/}\HOLFreeVar{Xs}\ensuremath{]} \HOLBoundVar{E\sp{\prime}}
